---
title: "Assignment 04"
author: "Kevin Hu"
date: "03/10/2019"
output: html_document
---
Required packages
```{r, message=FALSE, warning= FALSE}
library(gapminder)
library(tidyverse)
library(scales)
```

## Original data
The original data that will be used to reshape is the "gapminder" dataset in R:
```{r}
DT::datatable(gapminder)
```

## Exercise 1: Univariate Data Reshaping (Option 1)

### 1. Make a tibble with one row per year, and columns for life expectancy for two or more countries.

Codes and results:
```{r}
gap_wide <- gapminder %>%
  filter(continent == "Asia") %>% #filter the dataset to display only Asian countries
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)
DT::datatable(gap_wide)
```

comment: Life expectancy by countries in Asia in 1952 - 2007, is shown in the table above.

### 2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

Codes and results:
```{r, fig.width= 5, fig.height= 3}
cap <- "Comparison in life expectancy between China (red) and Taiwan (blue) over time (1952 - 2007).
Both series exhibited an increasing trend, but the gap between them was closing."
gap_wide %>% 
  ggplot(aes(x = year)) +
  labs(y = "Life Expectancy", caption = cap) +
  geom_point(aes(y = China), colour = "red") +
  geom_point(aes(y = Taiwan), colour = "blue") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0,face= "italic"))
```

### 3. Re-lengthen the data.

Codes and results:
```{r}
gap_wide %>% 
  pivot_longer(cols = -year,
               names_to = "country",
               values_to = "lifeExp") %>% 
  DT::datatable()
```

## Exercise 2 Multivariate Data Reshaping (Option 1)

### 1 Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.

Codes and results:
```{r}
gap_wide_2 <- gapminder %>%
  filter(country == "Canada" | country == "Brazil") %>% #filter the dataset to display only two countries
  pivot_wider(id_cols = year,
              names_from = country,
              names_sep = "_",
              values_from = c(lifeExp, gdpPercap))
DT::datatable(gap_wide_2)
```

### 2. Re-lengthen the data.

Codes and results:
```{r}
gap_wide_2 %>% 
  pivot_longer(cols = -year,
               names_to = c(".value","country"),
               names_sep = "_") %>% 
  DT::datatable()
```

## Exercise 3: Table Joins
load required datasets
```{r message=FALSE, warning=FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```
### 1. For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.

Note: In the email tibble, each email seems corresponding to a party instead of individual name. Firstly, I will use separate_rows to make each person gets their own row, then join the two tibbles by the name column using left join.

Codes and results:
```{r paged.print=TRUE}
email_name <- email %>% 
  separate_rows(guest, sep = ", ")
colnames(email_name)[1] <- "name" #change the column name in the email tibble to match the guest tibble
DT::datatable(guest_email <- guest %>% 
  left_join(email_name, by = "name") %>% 
  select(party, name, email,everything()))
```

### 2. Who do we have emails for, yet are not on the guestlist?

Note: This can be achieved by antijoining the guestlist to the email list

Codes and results:
```{r}
(newlist <- email_name %>% 
  anti_join(guest, by = "name"))
   
```

### 3. Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

Note: This can be achieved by just stacking the original guestlist and the list created in the previous step together using the bind_rows() function.

Codes and results:
```{r}
DT::datatable(bind_rows(newlist, guest_email))
```

