---
title: "Assignment 5"
author: "Kevin Hu"
date: "10/10/2019"
output: html_document
---
Load the required libraries
```{r message=FALSE, warning=FALSE}
library(gapminder)
library(tidyverse)
library(forcats)
```

## Exercise 1: Explain the value of the here::here package

  Using the here::here package will increase the reproducibility of your codes. The user can run your code without changing the absolute path of the root/project directory regardless what operation system the user is using. This is achieved by auto-detecting the root directory, then joining it (in a format that matches the user's platform) with sub-directories and file name that you specify when the here function is called.
  
## Exercise 2: Factor management

### Task 1 Drop Oceania in the gapminder dataset
The number of rows in the original data:
```{r}
nrow(gapminder)
```

The number of levels of _continent_ in the original data:
```{r}
levels(gapminder$continent)
nlevels(gapminder$continent)
```

Drop __Oceania__: Filter the Gapminder data to remove observations associated with the continent of Oceania
```{r}
DT::datatable(gap <- gapminder %>% 
  filter(continent != "Oceania") %>% 
  droplevels())
  
```

The number of rows after dropping:
```{r}
nrow(gap)
```

The number of levels of _continent_ after dropping:
```{r}
levels(gap$continent)
nlevels(gap$continent)
```

### Task 2 Reorder the levels of continent
Orignal order:
```{r}
levels(gap$continent)
```

Figure showing the orignal order:
```{r}
ggplot(gap, aes(continent, gdpPercap)) +
  geom_jitter() +
  labs(x = "Continent", y = "GDP per capita")
```

Reorder based on the standard deviation of GDP per capita:
```{r}
gap %>%
  ggplot() +
  geom_jitter(aes(fct_reorder(continent, gdpPercap, sd), gdpPercap)) +
  labs(x = "Continent", y = "GDP per capita")
```

## Exercise 3: File input/output (I/O)

### Summarize GDP per capita by continent
```{r}
gap_stat <- gapminder %>% 
  group_by(continent) %>% 
  summarise(mean_gdp = mean(gdpPercap),
            sd_gdp = sd(gdpPercap))
knitr::kable(gap_stat)
```

### Write the summary table as a csv file
```{r}
write_csv(gap_stat, here::here("gap_gdp_by_continent.csv"))
```

### Read the file and reorder the levels of continent based on the standard deviation of GDP per capita
```{r}
gap_import <- read_csv(here::here("gap_gdp_by_continent.csv"))
gap_import %>%
  ggplot() +
  geom_point(aes(fct_reorder(continent, sd_gdp), mean_gdp)) +
  labs(x = "Continent", y = "Average GDP per capita")
```

If you see the graph above, the newly created file survived the round trip of writing to file then reading back in.


